---
title: "Logistic Regression in Breast Cancer Patients"
author: "Carmen Canedo"
date: "10 August 2020"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# About
We will be looking at data collected on breast cancer patients in Wisconsin in 2020.

# Loading necessary libraries
These are the necessary libraries you will need to have loaded to follow along with the code below.

```{r, message=FALSE}
library(tidyverse)
```


# Loading in data
```{r}
# Reading in .csv
breast_cancer_original <- read.csv("Breast_cancer_data_Wisconsin_Stat302_2020.csv")
```

## Cleaning data
Renaming columns for clarity
```{r}
# Saving renamed columns into new data frame
breast_cancer <- breast_cancer_original %>% 
  rename(clump_thickness = ClumpThick) %>% 
  rename(size_uniformity = SizeUniforrm) %>% 
  rename(shape_uniformity = shapeUniform) %>% 
  rename(marginal_adhesion = Adhesion) %>% 
  rename(cell_size = cell.cize) %>% 
  rename(bare_nuclei = bareNuclei) %>% 
  rename(bland_cromatin = Cromatin) %>% 
  rename(normal_nucleoli = normalNuclei) %>% 
  rename(mitoses = Mitozis) %>% 
  rename(growth_class = Btype)
```

Normally, we would need to covert growth_class to a factor because its two values, 2 and 4, represent benign and malignant growths respectively. However, later in the code the `cor()` function requires that all columns are numeric. We will make those changes now.
```{r}
# Checking class
sapply(breast_cancer, class)

# List of column names
names <- c("ID", "clump_thickness", "size_uniformity", "shape_uniformity", "marginal_adhesion", "cell_size", "bare_nuclei", "bland_cromatin", "normal_nucleoli", "mitoses", "growth_class")

# Changing type to numeric using sapply and list of column names
breast_cancer[names] <- sapply(breast_cancer[names], as.numeric)

# Checking to make sure answer is correct
sapply(breast_cancer, class)
```


# Analysis
Let's take a look at the correlation matrix to get a general idea of the patterns in the data.
```{r}
bc_correlations <- cor(breast_cancer)

bc_correlations
```


We are going to use the backwards method to create our logistic model. We will first start with all variables included.
```{r}
# Creating model using all variables to determing growth_class
logistic <- glm(growth_class ~ ., data = breast_cancer, family = "binomial")

# Results of logistic model
summary(logistic)
```

## Interpretation
Using the information above, we can infer which variables have an effect on whether breast cancer patients in Wisconsin are likely to have benign versus malignant growths.

# Graphing
In order to graph the probabilities of the model below, we need to create a new data frame that has the probabilities.
```{r}
# Creating new data frame of probabilities
predicted_data <- data.frame(probability_of_growth_class = logistic$fitted.values, growth_class = breast_cancer$growth_class)

# Sorting from low to high
predicted_data <- predicted_data[order(predicted_data$probability_of_growth_class, decreasing = FALSE),]

# Ranking probability from low to high
predicted_data$rank <- 1:nrow(predicted_data)
```

## Creating graph

```{r}
ggplot(data = predicted_data, aes(x = rank, y = probability_of_growth_class)) +
  geom_point(aes(color = growth_class), alpha = 1, shape = 4, stoke = 2) +
  theme_set(theme_linedraw()) +
  labs(title = "Logistic Model of Breast Cancer Patients in Wisconsion",
       subtitle = "(1989-1991)",
       caption = "Data from Dr. William H. Wolberg, University of Wisconsin",
       x = "Index",
       y = "Predicted probability of begnign or malignant growth")
```

