---
title: "Logistic Regression in Breast Cancer Patients"
author: "Carmen Canedo"
date: "10 August 2020"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# About
We will be looking at data collected on breast cancer patients in Wisconsin in 2020.

# Loading necessary libraries
These are the necessary libraries you will need to have loaded to follow along with the code below.

```{r}
library(tidyverse)
```


# Loading in data
```{r}
breast_cancer_original <- read.csv("Breast_cancer_data_Wisconsin_Stat302_2020.csv")
```

## Cleaning data
Renaming columns for clarity
```{r}
breast_cancer <- breast_cancer_original %>% 
  rename(clump_thickness = ClumpThick) %>% 
  rename(size_uniformity = SizeUniforrm) %>% 
  rename(shape_uniformity = shapeUniform) %>% 
  rename(marginal_adhesion = Adhesion) %>% 
  rename(cell_size = cell.cize) %>% 
  rename(bare_nuclei = bareNuclei) %>% 
  rename(bland_cromatin = Cromatin) %>% 
  rename(normal_nucleoli = normalNuclei) %>% 
  rename(mitoses = Mitozis) %>% 
  rename(growth_class = Btype)
```

Next, we need to covert class to a factor because its two values, 2 and 4, represent benign and malignant growths respectively.
```{r}
breast_cancer$growth_class <- as.factor(breast_cancer$growth_class)
```


# Analysis